Antigenic phenotypes and phylogenetic metrics improve forecasts of seasonal influenza A/H3N2 evolution

### 2020-08-10 manuscript/main.tex
Add Pierre to the manuscript
Clarify figure captions especially for figure supplements
Expands figure supplement captions to be more understandable on their own and
cross-references each validation figure to its corresponding part of the
cross-validation schematic for simulated and natural populations. This latter
change should help readers distinguish between "validation" and "test" results.

  \figsupp[Composite model coefficients and distances to the future for models fit to simulated populations.]{
    A) Coefficients and B) distances are shown per validation timepoint and test timepoint as in Figure~\ref{fig:unadjusted_model_accuracy_and_coefficients_for_simulated_populations_controls}.}{\includegraphics[width=\textwidth]{figures/Figure_3-Figure_supplement_1.pdf}}\label{figsupp:unadjusted_composite_model_accuracy_and_coefficients_for_simulated_populations}

@@

\begin{figure}
  \includegraphics[width=\textwidth]{figures/Figure_4.pdf}
  \caption{
  Test of best model for simulated populations (true fitness) using 9 years previously unobserved test data and fixed model coefficients.
  A) The correlation of log estimated clade frequency fold change, $\log_{10}{\frac{\hat{x}(t + \Delta{t})}{x(t)}}$, and log observed clade frequency fold change, $\log_{10}{\frac{x(t + \Delta{t})}{x(t)}}$, shows the model's ability to capture clade-level dynamics without explicitly optimizing for clade frequency targets.
  B) The rank of the estimated best strain based on its distance to the future in the best model was in the top 20th percentile for 89\% of 18 timepoints, confirming that the model makes a good choice when forced to select a single representative strain for the future population.
  C) Absolute forecast error for clades shown in A by their initial frequency with a mean LOESS fit (solid black line) and 95\% confidence intervals (gray shading) based on 100 bootstraps.
  D) The correlation of all strains at all timepoints by the percentile rank of their observed and estimated distances to the future.
  The corresponding results for the naive model are shown in Figure~\ref{fig:test_of_best_model_for_simulated_populations}-Figure supplement~\ref{figsupp:test_of_naive_model_for_simulated_populations}.
  }
  \label{fig:test_of_best_model_for_simulated_populations}

  \figsupp[Validation of best model for simulated populations of H3N2-like viruses.]{
    A) The correlation of estimated and observed clade frequency fold changes shows the model's ability to capture clade-level dynamics without explicitly optimizing for clade frequency targets.
    B) The rank of the estimated best strain based on its distance to the future for 33 timepoints.
  The estimated best strain was in the top 20th percentile of observed closest strains for 100\% of timepoints, confirming that the model makes a good choice when forced to select a single representative strain for the future population.
  C) Absolute forecast error for clades shown in A by their initial frequency with a mean LOESS fit (solid black line) and 95\% confidence intervals (gray shading) based on 100 bootstraps.
  D) The correlation of all strains at all timepoints by the percentile rank of their observed and estimated distances to the future.
  The corresponding results for the naive model are shown in Figure~\ref{fig:test_of_best_model_for_simulated_populations}-Figure supplement~\ref{figsupp:validation_of_naive_model_for_simulated_populations}.}{\includegraphics[width=\textwidth]{figures/Figure_4-Figure_supplement_1.pdf}}\label{figsupp:validation_of_best_model_for_simulated_populations}

  \figsupp[Validation of naive model for simulated populations of H3N2-like viruses.]{
    Note that the naive model sets future frequencies to current frequencies such that there is no estimated fold change in frequencies for the first panel.}{\includegraphics[width=\textwidth]{figures/Figure_4-Figure_supplement_2.pdf}}\label{figsupp:validation_of_naive_model_for_simulated_populations}

  \figsupp[Test of naive model for simulated populations of H3N2-like viruses.]{
    Note that the naive model sets future frequencies to current frequencies such that there is no estimated fold change in frequencies for the first panel.}{\includegraphics[width=\textwidth]{figures/Figure_4-Figure_supplement_3.pdf}}\label{figsupp:test_of_naive_model_for_simulated_populations}

@@

\begin{figure}
  \includegraphics[width=\textwidth]{figures/Figure_7.pdf}
  \caption{
  Test of best model for natural populations of H3N2 viruses, the composite model of HI antigenic novelty and mutational load.
  A) The correlation of estimated and observed clade frequency fold changes shows the model's ability to capture clade-level dynamics without explicitly optimizing for clade frequency targets.
  B) The rank of the estimated best strain based on its distance to the future for eight timepoints.
  The estimated best strain was in the top 20th percentile of observed closest strains for 100\% of timepoints.
  C) Absolute forecast error for clades shown in A by their initial frequency with a mean LOESS fit (solid black line) and 95\% confidence intervals (gray shading) based on 100 bootstraps.
  D) The correlation of all strains at all timepoints by the percentile rank of their observed and estimated distances to the future.
  The corresponding results for the naive model are shown in Figure~\ref{fig:test_of_best_model_for_natural_populations}-Figure supplement~\ref{figsupp:test_of_naive_model_for_natural_populations}.
  }
  \label{fig:test_of_best_model_for_natural_populations}

  \figsupp[Validation of best model for natural populations of H3N2 viruses.]{
    Validation results for the composite model of mutational load and LBI.
    A) The correlation of estimated and observed clade frequency fold changes shows the model's ability to capture clade-level dynamics without explicitly optimizing for clade frequency targets.
    B) The rank of the estimated best strain based on its distance to the future for 23 timepoints.
    The estimated best strain was in the top 20th percentile of observed closest strains for 87\% of timepoints, confirming that the model makes a good choice when forced to select a single representative strain for the future population.
    C) Absolute forecast error for clades shown in A by their initial frequency with a mean LOESS fit (solid black line) and 95\% confidence intervals (gray shading) based on 100 bootstraps.
    D) The correlation of all strains at all timepoints by the percentile rank of their observed and estimated distances to the future.
    The corresponding results for the naive model are shown in Figure~\ref{fig:test_of_best_model_for_natural_populations}-Figure supplement~\ref{figsupp:validation_of_naive_model_for_natural_populations}.}{\includegraphics[width=\textwidth]{figures/Figure_7-Figure_supplement_1.pdf}}\label{figsupp:validation_of_best_model_for_natural_populations}

  \figsupp[Validation of naive model for natural populations of H3N2 viruses.]{
    Note that the naive model sets future frequencies to current frequencies such that there is no estimated fold change in frequencies for the first panel.}{\includegraphics[width=\textwidth]{figures/Figure_7-Figure_supplement_2.pdf}}\label{figsupp:validation_of_naive_model_for_natural_populations}

  \figsupp[Test of naive model for natural populations of H3N2 viruses.]{
    Note that the naive model sets future frequencies to current frequencies such that there is no estimated fold change in frequencies for the first panel.}{\includegraphics[width=\textwidth]{figures/Figure_7-Figure_supplement_3.pdf}}\label{figsupp:test_of_naive_model_for_natural_populations}

@@

We thank Jesse Bloom, Erick Matsen, Bing Brunton, Harmit Malik, Sidney Bell, Allison Black, Lola Arakaki, Duncan Ralph, and members of the Bedford lab for useful advice and discussions.
JH is a Graduate Research Fellow and is supported by the NIH grant NIAID F31AI140714.
The work done at the Crick Worldwide Influenza Centre was supported by the Francis Crick Institute receiving core funding from Cancer Research UK (FC001030), the Medical Research Council (FC001030) and the Wellcome Trust (FC001030).
SF, KN, NK, SW and HH were supported by the Ministry of Health, Labour and Welfare, Japan (10110400).
SW was supported by the Japan Agency for Medical Research and Development (JPfk0108118).
The Melbourne WHO Collaborating Centre for Reference and Research on Influenza is supported by the Australian Government Department of Health.
RAN is supported by NIAID R01 AI127893-01 and institutional core funding.
TB is a Pew Biomedical Scholar and is supported by NIH grants NIGMS R35 GM119774-01, NIAID U19 AI117891-01 and NIAID R01 AI127893-01.

@@

JH planned experiments, implemented the final forecasting framework, analyzed results, and wrote the manuscript.
JB, TR, XX, RK, DEW, LW, BE, RSD, JWM, SF, KN, NK, SW, HH, IB, and KS performed and provided data from serological assays.
RAN planned experiments and edited the manuscript.
TB planned experiments, implemented the initial forecasting framework, and edited the manuscript.

>>>>

  \figsupp[Composite model coefficients and distances to the future for models fit to simulated populations.]{
    Composite model coefficients and distances to the future for models fit to simulated populations.
    A) Coefficients and B) distances are shown per validation timepoint and test timepoint as in Figure~\ref{fig:unadjusted_model_accuracy_and_coefficients_for_simulated_populations_controls}.}{\includegraphics[width=\textwidth]{figures/Figure_3-Figure_supplement_1.pdf}}\label{figsupp:unadjusted_composite_model_accuracy_and_coefficients_for_simulated_populations}

@@

\begin{figure}
  \includegraphics[width=\textwidth]{figures/Figure_4.pdf}
  \caption{
  Test of best model for simulated populations (true fitness) using 9 years (18 timepoints) of previously unobserved test data and fixed model coefficients.
  These timepoints correspond to the open circles in Figure~\ref{fig:cross_validation_for_simulated_populations}.
  A) The correlation of log estimated clade frequency fold change, $\log_{10}{\frac{\hat{x}(t + \Delta{t})}{x(t)}}$, and log observed clade frequency fold change, $\log_{10}{\frac{x(t + \Delta{t})}{x(t)}}$, shows the model's ability to capture clade-level dynamics without explicitly optimizing for clade frequency targets.
  B) The rank of the estimated best strain based on its distance to the future in the best model was in the top 20th percentile for 89\% of 18 timepoints, confirming that the model makes a good choice when forced to select a single representative strain for the future population.
  C) Absolute forecast error for clades shown in A by their initial frequency with a mean LOESS fit (solid black line) and 95\% confidence intervals (gray shading) based on 100 bootstraps.
  D) The correlation of all strains at all timepoints by the percentile rank of their observed and estimated distances to the future.
  The corresponding results for the naive model are shown in Figure~\ref{fig:test_of_best_model_for_simulated_populations}-Figure supplement~\ref{figsupp:test_of_naive_model_for_simulated_populations}.
  }
  \label{fig:test_of_best_model_for_simulated_populations}

  \figsupp[Validation of best model for simulated populations of H3N2-like viruses.]{
    Validation of best model for simulated populations of H3N2-like viruses for 33 timepoints.
    These timepoints correspond to the closed circles in Figure~\ref{fig:cross_validation_for_simulated_populations}.
    A) The correlation of estimated and observed clade frequency fold changes shows the model's ability to capture clade-level dynamics without explicitly optimizing for clade frequency targets.
    B) The rank of the estimated best strain based on its distance to the future for 33 timepoints.
  The estimated best strain was in the top 20th percentile of observed closest strains for 100\% of timepoints, confirming that the model makes a good choice when forced to select a single representative strain for the future population.
  C) Absolute forecast error for clades shown in A by their initial frequency with a mean LOESS fit (solid black line) and 95\% confidence intervals (gray shading) based on 100 bootstraps.
  D) The correlation of all strains at all timepoints by the percentile rank of their observed and estimated distances to the future.
  The corresponding results for the naive model are shown in Figure~\ref{fig:test_of_best_model_for_simulated_populations}-Figure supplement~\ref{figsupp:validation_of_naive_model_for_simulated_populations}.}{\includegraphics[width=\textwidth]{figures/Figure_4-Figure_supplement_1.pdf}}\label{figsupp:validation_of_best_model_for_simulated_populations}

  \figsupp[Validation of naive model for simulated populations of H3N2-like viruses.]{
    Validation of naive model for simulated populations of H3N2-like viruses for 33 timepoints.
    These timepoints correspond to the closed circles in Figure~\ref{fig:cross_validation_for_simulated_populations}.
    Note that the naive model sets future frequencies to current frequencies such that there is no estimated fold change in frequencies for the first panel.}{\includegraphics[width=\textwidth]{figures/Figure_4-Figure_supplement_2.pdf}}\label{figsupp:validation_of_naive_model_for_simulated_populations}

  \figsupp[Test of naive model for simulated populations of H3N2-like viruses.]{
    Test of naive model for simulated populations of H3N2-like viruses for 18 timepoints.
    These timepoints correspond to the open circles in Figure~\ref{fig:cross_validation_for_simulated_populations}.
    Note that the naive model sets future frequencies to current frequencies such that there is no estimated fold change in frequencies for the first panel.}{\includegraphics[width=\textwidth]{figures/Figure_4-Figure_supplement_3.pdf}}\label{figsupp:test_of_naive_model_for_simulated_populations}

@@

\begin{figure}
  \includegraphics[width=\textwidth]{figures/Figure_7.pdf}
  \caption{
  Test of best model for natural populations of H3N2 viruses, the composite model of HI antigenic novelty and mutational load, across eight timepoints.
  These timepoints correspond to the open circles in Figure~\ref{fig:cross_validation_for_natural_populations}.
  A) The correlation of estimated and observed clade frequency fold changes shows the model's ability to capture clade-level dynamics without explicitly optimizing for clade frequency targets.
  B) The rank of the estimated best strain based on its distance to the future for eight timepoints.
  The estimated best strain was in the top 20th percentile of observed closest strains for 100\% of timepoints.
  C) Absolute forecast error for clades shown in A by their initial frequency with a mean LOESS fit (solid black line) and 95\% confidence intervals (gray shading) based on 100 bootstraps.
  D) The correlation of all strains at all timepoints by the percentile rank of their observed and estimated distances to the future.
  The corresponding results for the naive model are shown in Figure~\ref{fig:test_of_best_model_for_natural_populations}-Figure supplement~\ref{figsupp:test_of_naive_model_for_natural_populations}.
  }
  \label{fig:test_of_best_model_for_natural_populations}

  \figsupp[Validation of best model for natural populations of H3N2 viruses.]{
    Validation of best model for natural populations of H3N2 viruses for 23 timepoints using the composite model of mutational load and LBI.
    These timepoints correspond to the closed circles in Figure~\ref{fig:cross_validation_for_natural_populations}.
    A) The correlation of estimated and observed clade frequency fold changes shows the model's ability to capture clade-level dynamics without explicitly optimizing for clade frequency targets.
    B) The rank of the estimated best strain based on its distance to the future for 23 timepoints.
    The estimated best strain was in the top 20th percentile of observed closest strains for 87\% of timepoints, confirming that the model makes a good choice when forced to select a single representative strain for the future population.
    C) Absolute forecast error for clades shown in A by their initial frequency with a mean LOESS fit (solid black line) and 95\% confidence intervals (gray shading) based on 100 bootstraps.
    D) The correlation of all strains at all timepoints by the percentile rank of their observed and estimated distances to the future.
    The corresponding results for the naive model are shown in Figure~\ref{fig:test_of_best_model_for_natural_populations}-Figure supplement~\ref{figsupp:validation_of_naive_model_for_natural_populations}.}{\includegraphics[width=\textwidth]{figures/Figure_7-Figure_supplement_1.pdf}}\label{figsupp:validation_of_best_model_for_natural_populations}

  \figsupp[Validation of naive model for natural populations of H3N2 viruses.]{
    Validation of naive model for natural populations of H3N2 viruses for 23 timepoints.
    These timepoints correspond to the closed circles in Figure~\ref{fig:cross_validation_for_natural_populations}.
    Note that the naive model sets future frequencies to current frequencies such that there is no estimated fold change in frequencies for the first panel.}{\includegraphics[width=\textwidth]{figures/Figure_7-Figure_supplement_2.pdf}}\label{figsupp:validation_of_naive_model_for_natural_populations}

  \figsupp[Test of naive model for natural populations of H3N2 viruses.]{
    Test of naive model for natural populations of H3N2 viruses for eight timepoints.
    These timepoints correspond to the open circles in Figure~\ref{fig:cross_validation_for_natural_populations}.
    Note that the naive model sets future frequencies to current frequencies such that there is no estimated fold change in frequencies for the first panel.}{\includegraphics[width=\textwidth]{figures/Figure_7-Figure_supplement_3.pdf}}\label{figsupp:test_of_naive_model_for_natural_populations}

@@

We thank Jesse Bloom, Erick Matsen, Bing Brunton, Harmit Malik, Sidney Bell, Allison Black, Lola Arakaki, Duncan Ralph, and members of the Bedford lab for useful advice and discussions.
JH is a Graduate Research Fellow and is supported by the NIH grant NIAID F31AI140714.
The work done at the Crick Worldwide Influenza Centre was supported by the Francis Crick Institute receiving core funding from Cancer Research UK (FC001030), the Medical Research Council (FC001030) and the Wellcome Trust (FC001030).
SF, KN, NK, SW and HH were supported by the Ministry of Health, Labour and Welfare, Japan (10110400).
SW was supported by the Japan Agency for Medical Research and Development (JPfk0108118).
The Melbourne WHO Collaborating Centre for Reference and Research on Influenza is supported by the Australian Government Department of Health.
PB and RAN are supported by NIAID R01 AI127893-01 and institutional core funding.
TB is a Pew Biomedical Scholar and is supported by NIH grants NIGMS R35 GM119774-01, NIAID U19 AI117891-01 and NIAID R01 AI127893-01.

@@

JH planned experiments, implemented the final forecasting framework, analyzed results, and wrote the manuscript.
JB, TR, XX, RK, DEW, LW, BE, RSD, JWM, SF, KN, NK, SW, HH, IB, and KS performed and provided data from serological assays.
PB analyzed results.
RAN planned experiments and edited the manuscript.
TB planned experiments, implemented the initial forecasting framework, and edited the manuscript.
